---
title: "Hw2"
output:
  pdf_document: default
  html_document: default
date: '2023-04-01'
---

# the document is by R11323019 .

# 1.4.2 Let’s draw just N = 2, use optim function to get maximum likelihood estimates.

```{r}
library ( stats )
library ( tidyverse )

optimfunc <- function ( param ) {
  sig1 <-param [1]
  sig2 <-param [2]
  
  eplison1 = rnorm (2 ,0 , sig1 ^2) #N=2
  eplison2 = rnorm (2 ,0 , sig2 ^2)
  y = eplison1 + eplison2
  log( sig1 ^2+ sig2 ^2) +log (2*pi ) +(1 /(2*( sig1 ^2+ sig2 ^2) ) )*sum ( y ^2)
}

param <-c(2 ,1) # initial values
optimfunc.result <-optim ( param , optimfunc , method ="L-BFGS-B", hessian = TRUE )
optimfunc.result$par
```

# 3.2.1 Pick you favorite β1, β2 ≠ 0. Go back to your simulation in the first homework and re-simulate the model.


```{r}
## set value

mu0 <- 10
mu1 <- 9
var0 <- 1
var1 <- 2
var01 <- 0.6
beta1 <- 1
beta2 <- 2
```

```{r}
sigmav <-sqrt ( var0 ^2+ var1 ^2 -2* var01 )
rho <- var01 /( var0 * var1 )
sigma <-rbind (c( var0 ^2 , var01 ) ,c( var01 , var1 ^2) )

```



```{r}
library('dplyr')
library('data.table')
library(MASS)

simulation <- data.table(
  mvrnorm ( n =10^7 , mu = c(0,0) , Sigma = sigma),
  x1 =  rnorm(10^7, 1, var0),
  x2 =  rnorm(10^7, 1, var1)
)

colnames ( simulation ) <-c("eplison0", "eplison1", "x1", "x2")

```

```{r}
## create W0 and W1
# W0 = u0 + beta1*x1 + eplison0
# w1 = u1 + beta1*x1+  beta2*x2 eplison1

simulation[, w0:= mu0 + beta1*x1 +eplison0]
simulation[, w1:= mu1 + beta1*x1 + beta2*x2 + eplison1]
```



```{r}
simulation[, .(mean(w0),mean(w1),mean(x1),mean(x2))]
```

# 3.2.5 Create a column in your simulated data for the estimated propensity score using the derived formula above.

```{r}
# create True D
simulation[w1>w0, D:= 1]
simulation[w1<=w0, D:= 0]

#create p(x)
simulation[, propensity_score_formula := 1 - pnorm (( mu0 - mu1 - beta2*x2 )/ sigmav ,0 ,1)]


```


# 3.2.6 Use logit to estimate the propensity score
```{r}

logit <-glm(D~x2 , data = simulation , family = binomial ( link ="logit") )

simulation[, propensity_score_logit := predict(logit, type ="response")]


```

# 3.2.7 What is the correlation coefficient of the above two types of propensity scores?
```{r}
cor ( simulation$propensity_score_formula ,simulation$propensity_score_logit )
```

# 3.2.8 Use both types of propensity score to conduct IPW estimates.
```{r}
# create weight by formula
simulation[D==1, weight_formula := 1/propensity_score_formula]
simulation[D==0, weight_formula:= 1/(1-propensity_score_formula)]

# create weight by logit setimate
simulation[D==1, weight_logit := 1/propensity_score_logit]
simulation[D==0, weight_logit:= 1/(1-propensity_score_logit)]

#create actual wage
simulation[D==1, wage := w1]
simulation[D==0, wage := w0]
```

```{r}
ipw_formula <-lm( wage ~D, weights = weight_formula , data = simulation )
summary ( ipw_formula )
ipw_formula$coefficients

ipw_logit <-lm( wage~D, weights = weight_logit , data = simulation )
summary(ipw_logit)
ipw_logit$coefficients
```

# 3.2.9 regress wi on Xi.
```{r}
wage_to_D_OLS <-lm(wage ~D, data = simulation )
summary (wage_to_D_OLS)
wage_to_D_OLS$ coefficients
```

# 3.2.10 Now estimate by adding “control variables.
```{r}
wage_to_D_and_x2_OLS <-lm( wage~D+x2 , data = simulation )
summary ( wage_to_D_and_x2_OLS )
wage_to_D_and_x2_OLS$ coefficients
```












